---
title: 'Assignment #1'
author: "Diego Rios"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Setting of the working environment
```{r}
rm(list=ls())
getwd()
setwd("~/GitHub/QB2019_Rios/2.Worksheets/3.RStudio")
```

R as a calculator
```{r}
#Volume of a cube
l <- 5
l^3

#Area of a circle with r = 2
r <- 2
pi*r^2

#Length of the opposite side of a right-triangle
sin(pi/4)*sqrt(2)

#log (base e) of my favorite number
log(pi)
```

#Working with Vectors
##Basic Features of Vectors
```{r}
#First part
x <- rnorm(5, mean = 0, sd = 1)
w <- x*14
(x+w)/15
k <- 5*x
d <- c(w[1:3],k[1:4])
```

#Summary Statistics of Vectors
```{r}
v <- c(16.4,16,10.1,16.8,20.5,NA,20.2,13.1,24.8,20.2,25,20.5,30.5,31.4,27.1)
max(v,na.rm = T) #max
min(v,na.rm = T) #min
sum(v,na.rm = T) #sum
mean(v,na.rm = T) #mean
median(v,na.rm = T) #median
var(v,na.rm = T) #var
sd(v,na.rm = T)/sqrt(length(v)) # Standard Error of the mean
```

##Working with Matrices
```{r}
col1 <- rnorm(5, mean=8,sd=2)
col2 <- rnorm(5,mean=25,sd=10)
tarea1 <- cbind(col1,col2)
```
##Answer 1:
###The rnorm function generates random numbers according to a normal distribution function. The arguments it has specify the number of observations (n), the mean, and the standard deviation of the desired distribution.


```{r}
m <- read.table("data/matrix.txt")
m <- t(m)
dim(m)
```
##Answer 2:
###5 rows and 10 columns.

```{r}
#Indexing a Matrix
n<-m[-5,-3]
```


##Basic data visualization and statistical analysis
```{r}
meso <- read.table("data/zoop_nuts.txt",header = T)
meso.num <- as.matrix(meso[,c(-1,-2)])
pairs(meso.num)
cor1 <- cor(meso.num)
cor1
```
##Answer 3:
###All variables, except for CHLA, are highly and positively correlated. CHLA is weakly and negatively correlated with the remainder variables.

```{r}
library(psych)
cor3 <- corr.test(meso.num, method="pearson",adjust = "BH")
print(cor3, digits = 3)

cor4 <- corr.test(meso.num, method="spearman",adjust = "BH")
print(cor4, digits = 3)
```
##Answer 4
The direction and strength of the correlations don't change with the type of analysis used.
One should use non-parametric methods when the data doesn't follow a normal distribution. 
There doesn't seem to be evidence of false discovery rate given the high significance (p< 0.001) of the correlations. False discovery rate is important because one doesn't want to accept false hypotheses


Linear Regression
```{r}
fitreg <- lm(ZP~TN, data=meso)
summary(fitreg)
plot(meso$TN, meso$ZP, ylim = c(0,10), xlim=c(500,5000),
     xlab= expression(paste("Total Nitrogen (",mu,"g/L)")),
     ylab= "Zooplancton Biomass (mg/L)",las=1)
text(meso$TN, meso$ZP, meso$NUTS, pos=3, cex= 0.8)
newTN <- seq(min(meso$TN), max(meso$TN), 10)
regline <- predict(fitreg, newdata = data.frame(TN=newTN))
lines(newTN, regline)
conf95 <- predict(fitreg, newdata = data.frame(TN=newTN),
                interval = c("confidence"), level = 0.95, type= "response")
matlines(newTN, conf95[,c("lwr","upr")], type="l", lty = 2, lwd = 1, col = "black")
```

##Answer 5
Total nitrogen affects positively Zooplancton Biomass (p < 0.001, R^2 = 0.55)


##ANOVA
```{r}
NUTS <- factor(meso$NUTS, levels = c('L','M','H'))
zp.means <- tapply(meso$ZP, NUTS, mean)
sem <- function(x){
  sd(na.omit(x))/sqrt(length(na.omit(x)))
}
zp.sem <- tapply(meso$ZP, NUTS, sem)
bp <- barplot(zp.means, ylim = c(0, round(max(meso$ZP), digits = 0)),
              pch = 15, cex = 1.25, las = 1, cex.lab = 1.4, cex.axis = 1.25,
              xlab = "nutrient supply",
              ylab = "zooplankton biomass (mg/L)",
              names.arg = c("low","medium","high"))
arrows(x0 = bp, y0 = zp.means, y1 = zp.means - zp.sem, angle = 90,
       length = 0.1, lwd = 1)
arrows(x0 = bp, y0 = zp.means, y1 = zp.means + zp.sem, angle = 90,
       length = 0.1, lwd = 1)
fitanova <- aov(ZP ~ NUTS, data = meso)
summary(fitanova)
```

##Site by Species Matrix
```{r}
meso2 <- read.table("data/zoops.txt",header = T)
meso2 <- meso2[,c(-1,-2)]
site <- as.factor(seq(24))
meso2 <- cbind(site,meso2)
head(meso2)
```

